### 2015-07-07 narrative/intro.tex
new evaluation...
edits of intro

Avoiding impending bottlenecks in exascale scientific discovery will require
new research into managing and storing the large volumes of data that are 
produced by simulations and analyzed for months afterwards.
%
In this project we will demonstrate a cooperative approach for storing data
where the user and the storage system work together to achieve the best
possible utility and performance given data features, storage tier
characteristics, and current system state. We will investigate novel techniques
to facilitate efficient mapping of data objects
from the user space onto multiple storage tiers and enable
application-guided data reductions and transformations to address capacity and
bandwidth bottlenecks, while constraining errors to be within user provided
bounds.
% 
Our proposed research program is guided by the following overarching objectives.

{\em Address emerging storage architecture/technology landscape:}
We will address the associated I/O and storage challenges in the context of
current and emerging storage landscapes,
% and expedite insights into critical
%scientific processes, while demonstrating the validity of our approach in key
%DOE domains. 
%We will research techniques for creating and using a Scalable 
%Storage Software Infrastructure, and integrating services from the middleware 
%layer with Storage System capabilities to support 
and will develop novel strategies for distributing 
data both horizontally and vertically across the layers of these storage systems.  
Negotiation between these layers will be provided as a fundamental service, allowing users to
easily save ``the best'' amount of data, which is automatically placed throughout the
entire storage stack.

{\em Develop a science driven, system aware approach:}
%Our approach targets the expected characteristics of exascale storage hardware.
%The storage layer will be partitioned into multiple heterogeneous tiers with
%vastly different performance characteristics. The layers will be further
%differentiated by the constraints on capacity and data lifetime within each
%layer. Tape archival storage will still maintain data long term, but access to
%this data will be orders of magnitude slower than the next layer and naively
%accessing data from archival storage will greatly impact productivity.
%
Given that storage systems will be composed of multiple heterogeneous tiers with
vastly different performance characteristics and constraints on capacity and 
data lifetime they can support, our approach will enable researchers to 
place more knowledge into the SSIO system so that data can be recast from 
a sequence of bytes to information, which the system can then place and 
optimize the time to gain knowledge from exascale simulations.

@@

{\em Emphasize impacts on science:}
%The metric we are most interested in optimizing is time to knowledge.  
%Current approaches to addressing the I/O bottleneck fall into two broad categories:
%parallel file system approaches that optimize the throughput for an entire
%system, and I/O middleware approaches that optimize the performance of a single
%application. Both approaches have been successful to date, but are unlikely to
%overcome the remaining obstacles in reaching exascale. 
Instead of trying to optimize throughput, we will seek to reduce the time to knowledge.
This is the most significant metric for scientific applications where the desired outcome
is not storing data, but rather executing knowledge creation and extraction processes 
on the data. 
Moreover, we aim to perform the optimization not for a single
application, but rather for workloads within a multi-user multi-application
system environment.

The overarching idea of this proposal is that data from applications can be characterized by different
%levels of importance and different 
levels of importance over time, which then can be used to guide the analysis and visualization later to access the most important data and deliver some level of knowledge without waiting for access to the whole original dataset. Information about the importance of the data at the application level, information about resource status and performance of the multi-level storage system, and flexible data reorganization and reduction operations at all levels can be utilized to provide a quality of service and allow users to gain insight in a timely manner.

This proposal combines knowledge and expertise 
from a wide range of sources including ORNL applications, the ADIOS project,
the SNL Lightweight File Systems project and storage system knowledge, the metadata
management and storage systems knowledge of UCSC, and the middleware expertise
of Rutgers. We will integrate this diverse expertise to investigate how to make
a data, workflow, and application aware storage system.
This project brings together a team with strong expertise in I/O middleware (ORNL, Rutgers), 
file system (SNL, UCSC) and storage (UCSC), and connects and
coordinates these key storage components in a seamless fashion.

%
%\section{Scalable Storage Software Infrastructure}
%Our focus in this proposal is in Theme 2: {\em Scalable Storage Software Infrastructure} 
%We will focus on a state of the art SSIO solution to solve problems which researchers are 
%beginning to face and will make science on exascale systems
%problematic without new insights and solutions.
Our proposal will be responsive to Theme 2: {\em Scalable Storage Software Infrastructure} of the call, as summarize in Table~\ref{ab:call}.
\begin{table}[htbp]
\vspace{2ex}
\begin{center}
\begin{tabular}{ | p {0.2cm} | p{5cm} | p{7.5cm} | p{1.6cm} |}\hline
\multicolumn{2}{|c|}{\bf Call area} & \multicolumn{1}{|c|}{\bf Relevant S2E2 work} & \multicolumn{1}{|c|}{\bf See} \\\hline\hline
%\multicolumn{2}{|c|}{Collaboration} & 
\multicolumn{2}{|p {5.5cm}|}{Approaches to improve the ability of SSIO software to support checkpoint/restart;} & 
General piece we are doing for this... & 
\S\ref{sec:?}, \S\ref{sec:??}, \\\hline
1 & Data abstractions & What will we do & \S\ref{sec:?}, \S\ref{sec:?}, \S\ref{sec:?} \\\hline
2 & Mapping science data models onto hierarchical storage & What we are doing & \S\ref{sec:?}, \S\ref{sec:?} \\\hline
3 & Mechanisms for data movement across the storage hierarchy & what are we doing. & \S\ref{sec:?} \\\hline
4 &  Extend the interaction of IO middleware with the resource management system & What we are doing & \S\ref{sec:?}, \S\ref{sec:?}  \\\hline
5 & Design and implement of IO middleware architectures  & What we are doing & \S\ref{sec:?}, \S\ref{sec:?}, \S\ref{sec:?} \\
\hline
\end{tabular}
\end{center}
\label{tab:call}
\end{table}

@@

\subsection{Writing Perspective}
\label{subsec:sim-perspective}
The simulation scientist ideally should be able to
obtain an estimate of how much time will be required to write their data,
and how much storage space they will be able to get during the lifetime of
their simulations. Furthermore, they would like the ability to get an assurance
of Quality of Service such that they can react appropriately when the
expected bandwidth, for example, is less than what they desire. Ideally, the user
would specify a set of rules with which the system can make autonomic
decisions to determine what should be done.
%

One of the key aspects of our approach is to allow the users to define the
utility of their data; in other-words, the system should provide a
mechanism such that, for each data object, a user can specify pieces of the object
which have more utility (more importance).
There are many research challenges faced with
this approach including: 1) How do we make this simple enough so that users can
do this, 2) Can we define a model to allow the system to make
autonomic decisions to save the data which has the highest utility, and 3) Can
we allow the back-end storage system to interact with this model so that
it can optimize not just a single application but across all of the
applications.
 %
  In our simple scenario, a user wants to write 1~PB of data every hour
for checkpoint/restart, and they wish to write 500~TB of data every 15
minutes to save data for analysis and visualization. At this stage, they will ask the system to
write the data, and they would like to get an estimated time from which they can
then figure out, through a series of rules, whether they want to write out all
of the data, postpone the write until a later time, or write a reduced amount
of data using in situ reduction techniques, along with some code
container that allows for regenerating the data with a specified level of
accuracy.  Furthermore, they realize that the analysis data will be saved on
either the parallel file system, long-term tape storage, or on the campaign storage,
which is a medium-term storage
area with lower bandwidth than the parallel file system.
For each piece of data, the user will specify the lifetime so
that the data with the highest utility is kept around in fast storage for the longest
period of time, while the data with the least utility may be the first to be migrated to
lower layers of the storage hierarchy.

>>>>

Avoiding impending bottlenecks in exascale scientific discovery will require
new research into managing, storing and retreiving the large volumes of data that are 
produced by simulations and analyzed for months afterwards.
%
In this project we will demonstrate a cooperative approach for storing data
where the user and the storage system work together to optimize performance with
respect to user requirements, current system state, and characteristics of the tiered-storage system.
Our investigation will include the novel techniques for the efficient mapping of data objects
from the user space across a hierarchical tiered storage. These techniques will
include support for application-guided data reductions with user-defined error bounds
and transformations to optimize performance, capacity and minimize bandwidth bottlenecks.

% 
Our proposed research program is guided by the following overarching objectives:
\begin{itemize}
\item
{\em Address emerging storage architecture and technology landscape:}
We will address the associated I/O and storage challenges in the context of
current and emerging storage landscapes,
% and expedite insights into critical
%scientific processes, while demonstrating the validity of our approach in key
%DOE domains. 
%We will research techniques for creating and using a Scalable 
%Storage Software Infrastructure, and integrating services from the middleware 
%layer with Storage System capabilities to support 
and will develop novel strategies for distributing 
data both horizontally and vertically across the layers of these storage systems.  
Negotiation between these layers will be provided as a fundamental service, allowing users to
easily save ``the best'' amount of data, which is automatically placed throughout the
entire storage stack.
\item
{\em Develop a science driven, system aware approach:}
%Our approach targets the expected characteristics of exascale storage hardware.
%The storage layer will be partitioned into multiple heterogeneous tiers with
%vastly different performance characteristics. The layers will be further
%differentiated by the constraints on capacity and data lifetime within each
%layer. Tape archival storage will still maintain data long term, but access to
%this data will be orders of magnitude slower than the next layer and naively
%accessing data from archival storage will greatly impact productivity.
%
Given that storage systems will be composed of multiple heterogeneous tiers with
vastly different performance characteristics and constraints on capacity and 
supported data lifetime, our approach will enable researchers to 
place more knowledge into the SSIO system so that data can be recast from 
a sequence of bytes to information rich data. This information rich data can the
be used by the system to placement within the tiered storage that will optimize the time
to gain knowledge and insight from exascale simulations.

@@

\item
{\em Emphasize impacts on science:}
%The metric we are most interested in optimizing is time to knowledge.  
%Current approaches to addressing the I/O bottleneck fall into two broad categories:
%parallel file system approaches that optimize the throughput for an entire
%system, and I/O middleware approaches that optimize the performance of a single
%application. Both approaches have been successful to date, but are unlikely to
%overcome the remaining obstacles in reaching exascale. 
Instead of trying to optimize throughput, we will seek to reduce the time to knowledge.
As extracting knowledge is the only metric that matters for scientific applications, we will
focus not only upon data storage, but upon efficient execution of knowledge create and extraction
workflows.
Moreover, we aim to perform the optimization not only for a single
application, but rather for multiple workloads running on a multi-user, multi-application
system.
\end{itemize}

This proposal can be summarized as research into using the varying importance of simulation data over time to
ensure that the right portions of data, in the right format and resolution are available to scientists when they are needed.
Application-specific information about the importance of data will be used in conjuction with system- and resource-specific information
to optimize the scientists experience interacting with, and extracing knowledge and insight from simulation data.

%The overarching idea of this proposal is that data from applications can be characterized by different
%%levels of importance and different 
%levels of importance over time, which then can be used to guide the analysis and visualization later to access the most important data and deliver some level of knowledge without waiting for access to the whole original dataset. Information about the importance of the data at the application level, information about resource status and performance of the multi-level storage system, and flexible data reorganization and reduction operations at all levels can be utilized to provide a quality of service and allow users to gain insight in a timely manner.

This proposal combines knowledge and expertise 
from a wide range of sources including applications running at ORNL, the ADIOS project,
the SNL Lightweight File Systems project and storage system knowledge, the metadata
management and storage systems knowledge of UCSC, and the middleware expertise
of Rutgers. We will integrate this diverse expertise to investigate how to create
a data, workflow, and application aware storage system.
This project brings together a team with strong expertise in I/O middleware (ORNL, Rutgers),
file system (SNL, UCSC) and storage (UCSC), and connects and
coordinates these key storage components in a seamless fashion.

%
%\section{Scalable Storage Software Infrastructure}
%Our focus in this proposal is in Theme 2: {\em Scalable Storage Software Infrastructure} 
%We will focus on a state of the art SSIO solution to solve problems which researchers are 
%beginning to face and will make science on exascale systems
%problematic without new insights and solutions.
Our proposal will be responsive to Theme 2: {\em Scalable Storage Software Infrastructure} of the call, as summarize in Table~\ref{tab:call}.

\begin{table}[htbp]
\vspace{2ex}
\begin{center}
\begin{tabular}{ | p {0.2cm} | p{5cm} | p{7.5cm} | p{1.6cm} |}\hline
\multicolumn{2}{|c|}{\bf Call area} & \multicolumn{1}{|c|}{\bf Relevant S2E2 work} & \multicolumn{1}{|c|}{\bf See} \\\hline\hline
%\multicolumn{2}{|c|}{Collaboration} & 
\multicolumn{2}{|p {5.5cm}|}{Approaches to improve the ability of SSIO software to support checkpoint/restart;} & 
General piece we are doing for this... & 
\S\ref{sec:?}, \S\ref{sec:??}, \\\hline
1 & Data abstractions & What will we do & \S\ref{sec:?}, \S\ref{sec:?}, \S\ref{sec:?} \\\hline
2 & Mapping science data models onto hierarchical storage & What we are doing & \S\ref{sec:?}, \S\ref{sec:?} \\\hline
3 & Mechanisms for data movement across the storage hierarchy & what are we doing. & \S\ref{sec:?} \\\hline
4 &  Extend the interaction of IO middleware with the resource management system & What we are doing & \S\ref{sec:?}, \S\ref{sec:?}  \\\hline
5 & Design and implement of IO middleware architectures  & What we are doing & \S\ref{sec:?}, \S\ref{sec:?}, \S\ref{sec:?} \\
\hline
\end{tabular}
\end{center}
%\caption {WRITE A CAPTION FOR THIS TABLE}
\caption {}
\label{tab:call}
\end{table}

@@

\subsection{Writing Perspective (is there a better name for this section????)}
\label{subsec:sim-perspective}
The simulation scientist ideally should be able to
obtain an estimate of how much time will be required to write the data,
and how much storage space is available during the simulations lifetime.
Further, they would like Quality of Service assurances so that reactions
to varying bandwidth can be handled appropriately. Ideally, a set of
user-specified rules would allow the system to make autonomic decisions dynamically.
%

Central to our approach is giving users the ability to define the importance, or utility of their data.
This includes the ability to assign varying levels of importance to individual data objects in their entirety, or specific
subsets and pieces.
Research challenges include 1) an expressive and easy to use interface, 2) a model for the system supporting
autonomic decision for saving data using the user-defined importance and utility of data, and 3) the interface between
the system and the back-end storage across both single, and multiple running applications.

 %
  In our simple scenario, a user wants to write 1~PB of data every hour
for checkpoint/restart, and write 500~TB of data every 15
minutes for analysis and visualization.
The user would then specify a set of rules that governed the actual behaviour while the simulation is running.
If the system resources are available and the estimated write times are in order, the write can proceed.
Or, the write can be postponed to a later time-step, or data processed via in situ techniques
or a reduced set of data along with a regenerating code container at a given accuracy can be saved.
Data importance or utility will be used to place the various data on the proper resources
either the parallel file system, long-term tape storage, or on the campaign storage (medium-term storage with lower bandwidth).
The user specified lifetime for each data ensures that the data with the highest utility is preserved in fast storage
for the longest period of time, and the least utility are migrated to lower layers in the storage hierarchy.
